\chapter{Introduction} \label{ch:intro}

Factorization theory pursues a full understanding of how complex objects decompose into their simplest constituent parts.  
Depending on the algebraic structure in question, the difficulty of gaining such an understanding can vary wildly. 
Some objects can be broken down in exactly one way, while others exhibit more exotic behavior and are able to be broken down into several qualitatively different combinations of simpler parts.
Among our tasks are to test the bounds of this behavior, and to completely classify the circumstances under which it can occur.
In the present work, we bring our attention to a fairly new class of algebraic objects -- the titular ``power monoids" --  which possess many characteristics that make their study difficult, hence interesting.

\section{Motivation}
%%%%%%
%%%%%%

%The most elementary setting in which we study factorizations is the set $\ZZ$ of integers.
%It is well known (as the Fundamental Theorem of Arithmetic) that every integer (other than $-1$, $0$, and $1$) factors uniquely as a product of prime integers.
%For instance, $12$ can be written as $2\cdot 2\cdot 3$.
%Of course, 12 can also be written as either of the products $2\cdot 3\cdot 2$ or $(-3)\cdot 2\cdot (-2)$, but we consider these factorizations to be fundamentally the same.  
%This tells us that, in addition to identifying the prime factors involved, there should also be an equivalence of factorizations in play.
%Making these ideas rigorous is one of the challenges of extending this familiar example to a more general theory.


%There are many settings other than the integers in which it is reasonable to decompose elements into atoms. 
%However, most will not share the familiar unique factorization of the integers.  
%Historically, one of the greatest examples comes from the ring of integers of an algebraic number field; namely, $\ZZ[\sqrt{-5}]$.
%Consider $6$ as an element of this ring: $6=2\cdot3$, but we also have $6 = (1+\sqrt{-5})(1-\sqrt{-5})$.
%It is not a hard exercise to show that these two factorizations are not equivalent (meaning that $2$ and $3$ are not associate to $1\pm\sqrt{-5}$), so $6$ has more than one type of factorization into irreducibles.

%%For an example, one doesn't need to migrate too far from the integers: consider the subset of integers which can be obtained as a sum of any number of copies of $4$, $9$, or $11$.
%%The list of such integers begins: $\{0, 4, 8, 9, 11, 12, 13, 15, 16, 17, ... \}$.
%%Here, it is necessary to determine which integers will appear in the list (one can show that every integer after $15$ will appear), but we will be more interested in the questions of how many ways integers can be written just using $4$, $9$, and $11$.
%%As an example, $22$ can be written as $11+11$ or as $9+9+4$; so $22$ can be decomposed into two atoms or into three.
%%This is an instance of an element whose factorizations aren't unique, but a fairly tame one relative to the sort of phenomena that can arise within the study of non-unique factorizations.

%Much of the field has taken place in the setting of monoids which are not only commutative (satisfying $ab = ba$) but also cancellative, meaning that $ab = ac$ implies $b = c$.
%It is true that monoids of ideals in commutative rings and monoids of modules naturally give rise to examples of non-cancellative settings in which it is reasonable to study factorization properties. 
%Our goal here is to explore a relatively new class of monoids which are non-cancellative, exhibit many rich properties, and yet are rooted in a simple and natural combinatorial construction.

The motivation to study power monoids is drawn from two sources: factorization theory and arithmetic combinatorics.

Arithmetic combinatorics has historically involved sumsets in abelian groups (or even just in $\ZZ$).
It has been concerned with questions such as ``how large can a sumset be expected to be?'' which is resolved in the case of $\ZZ/p\ZZ$ by the Cauchy-Davenport inequality: for $A,B\subseteq \ZZ/p\ZZ$, $|A+B| \ge \min\{p,|A| + |B| -1\}$ \cite[Theorem 6.2]{grynkiewicz13}.
There is, of course, an accompanying inverse problem; namely, which pairs $A$ and $B$ satisfy $|A+B| = |A|+|B|-1$?
This is resolved by Vosper's Theorem, which says that the extremal pairs $A$ and $B$ are arithmetic progressions with the same difference \cite[Theorem 8.1]{grynkiewicz13}.
Though this is but a small slice of what the subject has to offer, it brings to our attention the potential for structural questions in arithmetic combinatorics and consequently highlights its surprising ideological closeness to parts of factorization theory.  

The broadest question in factorization theory is: ``how do elements decompose into atoms?''
Asking this question requires one to first specify (1) which elements, (2) which definition of ``atom'', and (3) which values of ``how''?
The first two points may be clear -- or should be after Chapter \ref{ch:fundamentals} -- but the third refers to what constitutes reasonable answers to the original question.  
Is it enough to specify that an element has more than one factorization? 
Or do we wish to distinguish factorizations by lengths, or by the types of atoms that they involve, or some further level of distinction?
Making these ideas rigorous is one of the challenges but also one of the predominantly interesting features of factorization theory.  

Once we have settled these parameters of the question, we can then think about various settings and how to characterize their factorization behavior.
Historically, one of the first classes of rings which drew the attention of many was that of rings of integers inside number fields.
In this setting, we find perhaps the most famous example of a ring which fails to have unique factorization, $\ZZ[\sqrt{-5}]$, as $2\cdot3 = 6 = (1+\sqrt{-5})(1-\sqrt{-5})$.
A significant effort, which drew techniques and ideas from analysis and geometry, went toward understanding the full range of behaviors exhibited by elements of rings of integers.
This eventually blossomed into the study of general Dedekind domains, Krull domains, and further generalizations.  

From this, we see how even a single class of examples can have a lasting impact on factorization theory, and on mathematics.
However, much of the field to date has focused on studying factorization in rings which are not only commutative but also \textit{cancellative}, meaning $ab = ac$ implies that $b=c$.
This is not to say that the subject has neglected non-cancellative settings altogether; monoids of modules or monoids of ideals in commutative rings are certainly well-studied.
However, these settings can involve heavy machinery and make it difficult to interact with other areas of mathematics.
On the other hand, power monoids are highly non-cancellative while being rooted in a simple and natural combinatorial construction.
We strive to open up the study of power monoids to bring attention to their own robust structure and also to explore their inevitable entanglement with combinatorics and number theory.  

%%%%%%
%%%%%%


\section{Plan and Main Results}
%%%%%%
%%%%%%

%In the remainder of this chapter, we will tell (some of) the story of factorization theory up to this point, by way of motivating the chapters to come.  
%We will also set some notation and conventions to be used.
In the remainder of this chapter, we will set some notation and conventions to be used hereafter.

Chapter \ref{ch:fundamentals} begins laying the foundation neessary to have a detailed discussion of factorizations.
Here we will give definitions and examples of properties which characterize the extend to which a monoid may fail to have uniqueness of factorization (the conditions UF, HF, FF, BF).  
The chapter ends with a comparison of the present framework to the existing body of literature, especially the work of D.D. Anderson et al.

Chapter \ref{ch:power monoids} introduces the main object of study: the power monoid, $\P_\fin(H)$.
We discuss the reduction, possible in many cases of interest, to the reduced power monoid $\P_\fun(H)$.
We determine that $\P_\fun(H)$ is atomic exactly when $H$ has no nontrivial idempotent elements or elements of order $2$ (Theorem \ref{th:atomicity}) and that $\P_\fun(H)$ is BF exactly when $H$ is torsion free (Theorem \ref{thm:BF-torsion}).

Chapter \ref{ch:minimal factorizations} returns to a discussion of factorization theory in general as we address the degeneracy of some of the usual notions in a non-cancellative setting.  
In response to this problem, we formulate the notion of a \textit{minimal factorization} and minimal versions of the usual factorization properties: UmF, HmF, FmF, BmF.
We then apply our new framework to power monoids to find that $\P_\fun(H)$ is BmF or FmF if and only if $\P_\fun(H)$ is atomic (Theorem \ref{BmF-char}); that $\P_\fun(H)$ is HmF if and only if $H$ is a group of order dividing $3$ (Theorem \ref{prop:HF-exp-3}); and that $\P_\fun(H)$ is UmF if and only if $H$ is trivial (Corollary \ref{cor:when-reduced-pm-is-minimally-factorial}).

In Chapter \ref{ch:partitions}, we begin to focus on the specific case of the reduced power monoid of $\NN$ with addition, which we refer to as $\PN$. 
In this setting we can leverage the linear ordering on $\NN$ to introduce the \textit{partition type} of a given factorization in $\PN$.
This is a new measure by which we may asses the degree to which an element fails to factor uniquely.
Indeed, we see that the intervals $[0,n]$ have factorizations of almost every possible partition type (Theorem \ref{thm:good types}).
This signifies a sharp dichotomy with the other elements of $\PN$, as any non-interval $X\in \PN$ satisfies $\mathsf{L}_\PN(X) \le \max(X)/2$ (Theorem \ref{thm:long factorization implies interval}).

Chapter \ref{ch:intervals} continues along the direction of quantifying the wild factorization behavior of intervals.
By constructing large families of atoms of $\PN$ and tying their growth rates to the growth of generalized Fibonacci numbers, we demonstrate that the number of factorizations of the interval $[0,n]$ grows exponentially with $n$.
In particular, for any $\e>0$, there is a constant $C$ such that, for sufficiently large $n$, $|\mathsf{Z}_\PN([0,n])| \ge C(\sqrt[4]{2}-\e)^n$ (Theorem \ref{thm:exponential intervals}).

Chapter \ref{ch:lattices} first establishes that $\PN$ and $\P_\fon(\NN^d)$ have, in a sense suitable for our purposes, identical factorization behavior for all $d>1$.
This affords us the opportunity to use higher-dimensional geometric intuition to attack problems in $\PN$ and leads us to develop some new methods for understanding factorizations in $\P_\fon(\NN^d)$.
This line of thought allows us to recover some known results for realizations of length sets in $\PN$; namely, that $\{n\} \in\mathcal{L}(\PN)$ for all $n\ge 2$ (Theorem \ref{thm:ind-atoms} or  \cite[Proposition 4.9]{fan-tringali18}) and that $\{2,n+1\}\in \mathcal{L}(\PN)$ for all $n\ge 2$ (Theorem \ref{thm:2n-length-set} or \cite[Proposition 4.10]{fan-tringali18}).
Finally, we push the methods developed to realize a new family of sets as sets of lengths in $\PN$; we show that $[2,m+2]\cup\{m+n+1\}\in \mathcal{L}(\PN)$ for all $n\ge 2$ and $m\ge 1$ (Theorem \ref{thm:int-point-construction}).

%%%%%%
%%%%%%
\section{Notation and Conventions} \label{subsec:generalities}

\begin{itemize}
	\item $\NN = \{0,1,2,3,\dots\}$, $\ZZ = \{0,\pm 1, \pm 2,\dots\}$, and $\RR$ denote the sets of natural numbers, integers, and real numbers, respectively.
	Here we adopt the French convention that $0\in \NN$ so that $\NN$ is a monoid.
	
	\item In general, unless otherwise specified, lowercase letters ($a$, $b$, $x$, $y$, etc.) will usually refer to elements of a monoid; 
	ordinary uppercase letters to sets and subsets ($A$, $B$, $X$, $Y$, etc.); 
	script or calligraphic uppercase letters ($\A$, $\C$, $\F$, $\P$) to collections of subsets; 
	math fraktur letters ($\aa$, $\bb$, etc.) to words in a free monoid over a given generating set 
	(see Section \ref{subsec:factorizations} for more details on free monoids).
	Specifically, $i$, $j$, $k$, $\ell$, $m$, and $n$ will usually stand for non-negative integers. 
	$H$ will stand for a monoid, $G$ for a group, and $R$ for a ring.
	
	\item $X\subseteq Y$ will mean that $X$ is a subset of $Y$; $X\subsetneqq Y$ will mean that $X\subseteq Y$ but $X\neq Y$; $K\le H$ will mean that $K$ is a submonoid of $H$.
	
	\item For a subset $S\subseteq \RR$ and $k\in \NN$, we set the notation $S_{>k} := \{n\in S: n>k\}$ and $S_{\ge k} := \{n\in S: n\ge k\}$.
	
	\item For $a, b\in \RR\cup\{\infty\}$, $[ a,b ] = \{ n\in \ZZ : a \le n \le b \}$ shall denote the (integer) interval from $a$ to $b$.
	
	\item For a real number $x$, $\lfloor x \rfloor := \max\{n\in \ZZ: n \le x\}$ (read \textit{floor of} $x$) is the greatest integer less than $x$.
	Similarly, $\lceil x \rceil := \min\{n\in \ZZ: n\ge x\}$ (read \textit{ceiling of} $x$) is the least integer greater than $x$.
	
	\item If $S$ is a set and $\E$ is an equivalence relation on $S$, the equivalence class of some $x\in S$ shall be denoted by $[x]_\E$.
	The subscript may be removed in situations where the implied equivalence is clear.
	
	\item Fix an integer $m>1$.  When $m$ is understood from context, $\bar{a} \in \ZZ/m\ZZ$ will denote the residue class of $a$ modulo $m$.
	We will write $a\equiv b \mod m$ if $\bar{a} = \bar{b}$.
	
	\item On occasion, we will wish to lift residue classes to elements of $\NN$.  
	If $x\in \ZZ/m\ZZ$ and $m$ is well understood from context, we set $\hat{x} := \min(x\cap \NN) \in [0,m-1]$.
	By a similar token, for a subset $X\subseteq \ZZ/m\ZZ$, we set $\hat{X} := \{\hat{x}: x\in X\}$.
	
	\item In general, we will refer to the operation of a not necessarily commutative monoid $(H,\cdot)$ as ``multiplication''.
	We will often adopt the practice of suppressing the symbol ``$\cdot$'' when no confusion is likely to arise; that is, we will write $xy$ instead of $x\cdot y$.
	
	\item The identity of a multiplicative monoid $(H,\cdot)$ will be called $1$; the identity of an additive monoid $(H,+)$ will be called $0$.
	
	\item For a $k\in \NN$ and elements $x_i$ of some set for $i\in [1,k]$, we will often write the family $X:=\{x_i: i\in [1,k]\}$ in ``long form'' as $x_1,\dots,x_k$.
	On the other hand, we may refer to its elements simply ``the $x_i$'' or $\{x_i\}_i$ for brevity.
	If $X$ is a subset of a monoid $(H,\cdot)$, we will write the ordered product of the elements of $X$ as $x_1\cdots x_k$.
	In particular, if $k$ happens to be $0$, this product is empty and we set the convention $x_1\cdots x_k = 1$ to avoid any confusion or ill-definedness.
\end{itemize}




%%%%%%
%%%%%%
